---
title: "Statistics_project"
author: "Seija Tuomi"
date: "22 May 2019"
output: html_document
---


library(tidyverse)

##Comparison Experiment

##NULL Hypothesis is - There is no difference between KCl and K2SO4 extraction of N species from any soil.  


```{r}
data<-read_csv("~/research.tuo/data_school_mine/stats/Comparison_Experiment_KCl_V_K2SO4.csv")
library(ggplot2)
```


#Import data and create new column for Treatment
#Create unique identifier by adding "value" of columnd 7, 9, 11 to make RepID column
#Create a new column for DON (Dissolved organic carbon) which is Total N (inorg + org) analysed - sum of N (inorganic) species analysed.

```{r}
K1<-read.csv("~/research.tuo/data_school_mine/stats/Comparison_Experiment_KCl_V_K2SO4.csv")
K1$Treatment<-rep(c("Urea","Water"), each=12, times=3)
K1$RepID<-with(K1,interaction(Treatment, Extractant, Soil))
K1$DON<-K1$TN_N_analy - K1$Sum_species
```

#Re-structure data
 
```{r}
K2<-gather(K1, key=N_type,value=amount, c(7,9,11))
glimpse(K2)
```

#Visualising Data
#Original units were Urea mg/L analysed but these plots were hard interpret

```{r}
ggplot((K2), aes(x=Soil, y=amount, colour=Extractant))+
  geom_boxplot()+
  facet_wrap(~ N_type)+
  geom_hline(yintercept = 4.66)
```

#ammonia in 3 soils
```{r}
ggplot(K1,aes(Soil, y=NH3N_adjusted, colour=Extractant))+ 
  geom_boxplot()
```

#nitrate in 3 soils
```{r}
ggplot(K1,aes(Soil, y=NO3_N_adjusted, colour=Extractant))+
  geom_boxplot()
```

#Urea in 3 soils
```{r}
ggplot(K1,aes(Soil, y=Urea_N_adjusted, colour=Extractant))+
  geom_boxplot()
```

#TN in 3 soils
```{r}
ggplot(K1,aes(Soil, y=TN_N_analy, colour=Extractant))+
  geom_boxplot()+
  geom_hline(yintercept = 4.66)
```

#(TN analysed) - (Sum_species)
```{r}
ggplot(K1,aes(Soil, y=DON, colour=Extractant))+
  geom_boxplot()
```

#DON = TN - NH3_N - NO3_N - Urea_N
```{r}
ggplot(K1,aes(Soil, y=Sum_species, colour=Extractant))+
  geom_boxplot()
```

#Why do we use log plots?

#There are two main reasons to use logarithmic scales in charts and graphs. The first is to respond to #skewness towards large values; i.e., cases in which one or a few points are much larger than the bulk #of the data. The second is to show percent change or multiplicative factors.
#Examples

```{r}
ggplot(K1, aes(Soil, log(Urea_N_adjusted), colour=Extractant))+geom_boxplot()+
  facet_wrap(~Treatment)
```
```{r}
ggplot(K1, aes(Treatment, log(Urea_N_adjusted), colour=Extractant))+geom_boxplot()+
  facet_wrap(~Soil)
```

##Remove water from data

##Marginal Effects
```{r}
K3<-subset(K1, Extractant != "H2O")
```

##Using log RE going up (homogenaity of variation)
#lmer is a linear mixed effects model
```{r}
library(lmerTest)
library(emmeans)
lm1<-lmer(log(NH3N_adjusted)~Extractant+Treatment+Soil +(1|RepID), data=K3)
```

```{r}
anova(lm1)
summary(lm1)
```
#The F-statistic is simply a ratio of two variances. Variances are a measure of dispersion, or how far #the data are scattered from the mean. Larger values represent greater dispersion.
#Variance is the square of the standard deviation.
# The t-value measures the size of the difference relative to the variation in your sample data.

## emmeans(lm, pairwise~Extractant,type="response")
```{r}
emmeans(lm1, pairwise~Extractant,type="response")
```
#Plotting lm1
```{r}
plot(lm1)
```




##Factorial Model

lm2<-lmer(log(NH3N_adjusted)~Extractant*Soil +(1|RepID), data=K3)
anova(lm2)
summary(lm2)
emmeans(lm2,pairwise~Extractant|Soil, type="response")
emmeans(lm2,revpairwise~Extractant|Soil, type="response")

plot(lm2)

 
